Direct Entropy Measurement in a Mesoscopic Quantum System

### 2018-02-09 paper/draft.tex
change some figures. remove most of a paragraph.

%%%% outstanding mysteries %%%%%
% error bar on 0-1 peak height (mention in text)
% can I cite someone as to why \Delta_ST changes with field
% dS or \partial S?
% why does \partial S/ \partial N

@@

%\begin{abstract}
%
%Measuring the entropy of an electronic state is a powerful tool for identifying its underlying microscopic character.  Such measurements are typically based on bulk properties, such as heat capacity, that are straightforward to observe in macroscopic samples but exceedingly difficult to access in mesoscopic systems that may consist of just a few electrons. Taking advantage of a well-known Maxwell relation, we realize a protocol for entropy-to-charge conversation in a gate-defined GaAs quantum dot that enables an entropy measurement of the first three quantum states in to the dot. The entropy of a single spin ($k_B \ln{2}$) is measured within 8\% accuracy, as is the entropy arising at the magnetic field-driven singlet-triplet crossing for two electrons.
%
%\end{abstract}

@@

%%%%%%% intro material %%%%%%%%%
\textbf{The thermodynamic properties of electronic systems can offer important insights into the nature of their ground states, and probe exotic quasiparticles that may emerge due to interactions or non-trivial topology.  Systems that are difficult to clearly identify through standard conductance measurements may be studied more in depth if a thermodynamic measurement can be made. For example, the purported non-Abelian exchange statistics of Moore-Read quasiparticles in the $\nu = \frac{5}{2}$ fractional quantum Hall state, or of Majorana quasiparticles in a topological superconductor, are exceedingly difficult to identify from conductance signatures. However, an entropy measurement could clearly distinguish Abelian from non-Abelian quasiparticles \cite{Cooper2009, Smirnov2015}.  Similarly, the two-channel Kondo state that is believed to arise in carefully tuned GaAs devices has so far been identified through a particular temperature dependence in the device conductance \cite{Potok2007}. A more direct test for the two-channel Kondo state would be a confirmation of its entropy ($\frac{1}{2} k_B \ln{2}$), that should remain down to arbitrarily low temperatures \cite{Alkurtass2016}.}

%%% figure 1 %%%
\begin{figure}[!]
        \includegraphics[width=1.0\columnwidth]{../figures/figure_1_no-annotation.pdf}
        \caption{\label{fig:fig1}(a) Scanning electron micrograph of a device similar to the one measured, showing electrostatic gates (blue) used to define a circuit in the 2D electron gas. Grey gates are unused (grounded). Squares show the locations of ohmic contacts to the 2DEG. An electron reservoir is formed between the parallel gates running vertically down the image. The reservoir can be heated by driving current through the quantum point contact on the left. A 200nm diameter quantum dot is coupled to the right side of the channel. Electrons tunneling between the reservoir and dot are measured with a capacitively-coupled charge sensor. (b) and (c) Charge sensor signal for a single transition from $N \rightarrow N+1$ electrons at two temperatures ($T_{Red} > T_{Blue}$) showing the two possible cases for $\frac{\partial \mu}{\partial T}$.}
\end{figure}

@@

In order to accomplish this entropy to charge conversion, we make use of the Maxwell relation
%
\begin{align}
\label{eqn:max}
        \left(\frac{\partial \mu}{\partial T}\right)_{p,N} &= -\left(\frac{\partial S}{\partial N}\right)_{p,T}
\end{align}
%
that connects changes in entropy, particle number, and temperature ($S$, $N$, and $T$, respectively) to changes in the chemical potential, $\mu$, a quantity that is simple to measure and control. The fixed pressure condition of Eq.~\ref{eqn:max} is effectively met by working well below the Fermi temperature, $T_F \sim$\SI{5}{\kelvin}, where degeneracy pressure dominates \cite{Landau1969}.

The Maxwell relation in Eq.~\ref{eqn:max} forms the basis of two exciting theoretical proposals to measure the entropy derivative, $\frac{\partial S}{\partial N}$, of the $\nu = \frac{5}{2}$ state by detecting changes in the quasiparticle chemical potential with temperature \cite{Cooper2009,Ben-Shach2013}.  Our work builds on these proposals to measure the entropy of the few-electron ground states of a quantum dot. Unlike the $\nu = \frac{5}{2}$ quasiparticles described in the theoretical work, the few-electron quantum dot states studied here have an entropy that is well-understood from the spin degree of freedom \cite{Tarucha1996, Ciorga2000, Duncan2000, Lindemann2002, Potok2003, Hofmann2016}. Our experiment demonstrates a direct entropy measurement is possible on a few-particle system. In the future, the method can be utilized as a probe into the nature of more exotic systems, such as topologically non-trivial Majorana and other non-Abelian states.

The mesoscopic circuit shown in Fig.~\ref{fig:fig1}a uses electrostatic gates to realize an electron reservoir (central region; hereafter ``reservoir") in thermal and electrochemical equilibrium with a few-electron quantum dot (``dot") coupled to its right side.  The occupation of the dot is measured using an adjacent quantum point contact (``charge sensor")\cite{Staring2007, Frolov2009, Thierschmann2015}.  An additional electron is added to an $N$ electron dot when the chemical potential, $\mu_{N+1}$, drops below the Fermi level of the reservoir, $E_F$. Figure 1b illustrates a such a transition \textemdash a step in the charge sensor conductance as a function of plunger gate voltage, $V_G$, which tunes the chemical potential on the dot, $\mu_{dot}$ \textemdash thermally broadened by the reservoir temperature, $T_{res}$.  The gate voltage corresponding to midpoint of the transition, $V_0$, marks the chemical potential at which the probabilities of finding $N$ or $N+1$ electrons on the dot are equal.

The relation between entropy and a temperature-induced shift in $\mu_{N+1}$ (with respect to $E_F$) can be understood in the language of detailed balance.  The $N+1^{\rm st}$ electron tunnels back and forth between an $N$-electron dot and a reservoir as long as there are available states in both dot and reservoir; true when $\mu_{N+1}$ is within thermal broadening of $E_F$.  The tunnel rates into and out of the dot, $\Gamma_{in}=\Gamma_{N\rightarrow N+1}$ and $\Gamma_{out}=\Gamma_{N+1\rightarrow N}$, depend on degeneracy of the $N$ and $N+1$ ground states.  Consider the filling of the first electron state, which has a single, unpaired spin-$\frac{1}{2}$ and, therefore, a degeneracy of 2. An electron tunneling into the dot has both spin-up and -down states available.  Tunneling out, the electron (whether spin-up or -down) must tunnel into reservoir states with the same spin, that is, only half of the reservoir states are available. As a result, $\Gamma_{in} = 2\Gamma_{out}$ when $\mu_{N+1}=E_F$. Extending this logic to a general case gives the result derived from detailed balance \cite{Gustavsson2009}, 
%
\begin{align}
	\Gamma_{in} &=  g_{N+1} \Gamma f(E_F - \mu_{N+1}) \nonumber \\
	\Gamma_{out} &= g_{N} \Gamma [1 - f(E_F - \mu_{N+1})] \label{eqn:rates}
\end{align}
%
where $\Gamma$ is a coupling constant, $f$ is the Fermi function, and $g_{N(N+1)}$ is the degeneracy of the $N(N+1)$ electron state.

The midpoint of the transition between the $N$ and $N+1$ electron states ($V_0$), as measured by the charge sensor, occurs when $\Gamma_{in} = \Gamma_{out}$ and the dot is occupied half the time. From Eq.~\ref{eqn:rates}, $\Gamma_{in} = \Gamma_{out}$ corresponds to $\mu_{N+1} = E_F$ when $g_{N}=g_{N+1}$. When $g_{N}\neq g_{N+1}$, $\mu_{N+1}$ shifts away from $E_F$ by an amount that depends on the broadening of the Fermi function, and therefore on $T$.   A measurement of how $V_0$ shifts with T allows the ratio $g_{N+1}/g_{N}$ to be determined, which itself is related to $dS_{N\rightarrow N+1}$ through the Boltzmann entropy, $S_{N}=k_{B} \ln{g_N}$.

Looking back to Eq.~\ref{eqn:max}, the shift in $\mu_{N+1}$, and therefore $V_0$, with temperature is proportional to the entropy difference between the $N$ to $N+1$ ground states ($\partial S_{N\rightarrow N+1}$ for $\partial N=1$).  This shift, $\partial V_0$ as a function of $\partial T$, is the quantity probed directly in the experiment described below. From these measurements the entropies of the first three energy levels in the quantum dot are built up, one by one. While the relationship between degeneracies and tunnel rates was recently explored in a time-resolved measurement \cite{Hofmann2016}, the thermal technique presented in this work extends the measurement of entropy (degeneracy) to a wider set of applications, where the tunneling process may not be convenient, or even possible, to observe in real-time.

@@

%%% figure 2 %%%
\begin{figure}[!]
        \includegraphics[width=1.0\columnwidth]{../figures/figure_2.pdf}
        \caption{\label{fig:fig2}(a) Charge sensor data for $N=0 \rightarrow 1$ at two temperatures set by DC current through the QPC heater. (b) Lock-in measurement of $dg_{sens}$ with $dT \sim 60mK$. Fits to $dg_{sens}$ are shown with $\delta$ as a free parameter (solid) and $\delta=0$ (dashed) (c) Transition width, $\Theta$, as a function of mixing chamber temperature. The highlighted region shows roughly where tunnel rates are determined only by temperature broadening. The lever arm $\alpha$ is calculated by fitting a straight line to this region. (d) $\Theta$ as a function of DC current through the QPC heater. Fit is used to convert between $I_{heat}$ and $dT$, with $R_{QPC} = $ \SI{20}{\kilo\ohm} (e) Change in entropy as a function of AC current through the QPC heater. Top axis shows corresponding $dT$. Right axis shows shift in $\partial V_0$ per unit temperature.}
\end{figure}

@@

%%% figs 2a and b %%%
The addition of the first electron to the dot was marked by a decrease in $g_{sens}$ that is well fit by the expected form for a thermally-broadened transition (Fig.~\ref{fig:fig2}a):
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_0(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_0(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $G_0$ quantifies the sensor sensitivity, $\Theta = \frac{k_B T}{\alpha e}$ is the thermal broadening expressed in units of gate voltage, $\alpha$ is the lever arm that converts $V_G$ to $\mu_{dot}$, and $G_1$ is determined by the cross capacitance between the charge sensor and plunger gate. The dependence of $V_0$ on $\Theta$ captures any shift in $\mu$ due to a change in entropy across the transition.

The transition width, $\Theta$, extracted from fits to Eq.~\ref{eqn:g} follows the mixing chamber temperature, $T_{MC}$, to \SI{<100}{\milli\kelvin} (Fig.~\ref{fig:fig2}b). The assumption of thermally broadened transitions greatly simplifies the data analysis, so $T_{MC}$ is fixed at \SI{100}{\milli\kelvin} for most of this work with Joule heating giving reservoir temperatures up to \SI{250}{\milli\kelvin}. Measurements of $\partial S$ are made at $dT_{res} = $ \SI{60}{\milli\kelvin} such that no orbital levels above the ground state were excited.

%%% figs 2c and d%%%
The data in Fig.~\ref{fig:fig2}c, and corresponding fits, offer an example of how $dS_{0\rightarrow 1}$ was measured across the $0 \rightarrow 1$ transition. A comparison of the 100 and 160 mK data in Fig.~\ref{fig:fig2}a shows that $g_{sens}$ drops as the temperature rises before the midpoint of the transition ($V_G-V_0<0$), whereas the conductance is higher in the warmer curve for $V_G-V_0>0$.   Oscillating the reservoir temperature by $dT_{res}$ using an AC current, $I^{AC}_{Heat}$, measuring oscillations in $g_{sens}$ with a lock in amplifier (see Methods) yields the characteristic peak-dip structure seen in Fig.~\ref{fig:fig2}c.  The expected lineshape of such a curve is given by the derivative of Eq.~\ref{eqn:g} with respect to temperature, expressed as $dg_{sens}$ due to a temperature change $dT$:
%
\begin{align}
\label{eqn:dg}
        dg_{sens}(V_G, \Theta) &\propto dT \left[ \frac{V_G-V_{0}(\Theta)}{2\Theta} +\delta \right]\times \\
        				      &\quad\cosh^{-2}\left(\frac{V_G-V_{0}(\Theta)}{2\Theta}\right) \nonumber
\end{align}
%
where $\delta=\frac{\partial V_0}{\partial \Theta}$. Using the lever arm to convert $V_0$ to chemical potential, the parameter $\delta$ is expressed as the change in entropy scaled by Boltzmann's constant:
%
\begin{align}
\label{eqn:delta}
        \delta &= \frac{\partial V_0}{\partial \Theta} = 
        \frac{1}{k_B} \frac{\partial \mu}{\partial T} = 
        -\frac{1}{k_B} \partial S_{N\rightarrow N+1}
\end{align}
%

From Eqs.~\ref{eqn:dg} and \ref{eqn:delta}, $dg_{sens}(V_G)$ would be perfectly antisymmetric around $V_0$ if $\partial S_{N\rightarrow N+1}$ were zero.  Fits to the data in Fig.~\ref{fig:fig2}c illustrate clearly that $\partial S_{0\rightarrow 1}=0$ is inconsistent with the data;  the best fit is found for $\partial S_{0\rightarrow 1}=0.71\times k_B$.  Indeed, this is what would be expected for the $0 \rightarrow 1$ electron transition at zero magnetic field: $S_0$ for the empty state should be $k_B \ln{1}=0$ whereas for the spin degenerate 1-electron state one expects $S_1=k_B\ln{2}$, giving $dS_{0\rightarrow 1}\equiv S_1 = S_0 =k_B\ln{2}$. Fig.~\ref{fig:fig2}e shows that $dS$,  determined from fits to data analogous to Fig.~\ref{fig:fig2}c, remains constant over a range of $dT$ ($I^{AC}_{Heat}$).  Although the extracted $\delta$ can be converted to gate-voltage-per-temperature units using a value for $\alpha$ measured separately (right axis of Fig.~\ref{fig:fig2}e), the entropy itself, $dS$, is determined directly from the fit to $g_{sens}$ without any additional calibration.

@@

Confirmation that the measured $dS$ derives from spin degeneracy can be seen in its evolution with in-plane magnetic field, $B_\parallel$. Figure \ref{fig:fig3}a compares $dS(B_\parallel)$ for the 0-1 and 2-3 electron transitions.  The 2-electron ground state is a spin singlet, while the 3-electron ground state adds an additional unpaired electron, so both $dS_{0 \rightarrow 1}$ and $dS_{2 \rightarrow 3}$ correspond to transitions from total spin zero $(S_0=S_2=0)$ to total spin one-half.One expects identical field-dependent behaviors for $dS_{0\rightarrow 1}$ and $dS_{2\rightarrow 3}$ as long as $g \mu_{B} B_\parallel$ is less than the excited state energies.

@@

The entropy measurement yields $\partial S(B=0) = (0.933 \pm 000) k_B \ln{2}$, which agrees quantitatively with the expected $k_{B} \ln{2}$, collapsing to zero at high field where spin degeneracy is broken.  At a qualitative level, Figs.~\ref{fig:fig3}b and c show the high-field crossover to a perfectly antisymmetric lineshape for $dg_{sens}(V_G)$; clearly indicating $\delta$ from Eqn.~\ref{eqn:dg} going to zero, corresponding to no entropy change between the two charge states. Using a $T=T_{res}$ of \SI{130}{\milli\kelvin} that is midway between the low ($T_{MC}$) and high ($T_{MC}+dT$) points of the temperature oscillation yields $g=-0.39$ or $g=-0.40$ for the $0\rightarrow 1$ and $2\rightarrow 3$ transitions respectively. This is consistent with the value $g=-0.44$ expected for GaAs and measured via bias spectroscopy in Fig.~\ref{fig:fig3}e.

@@

The field-dependent entropy measurement for the 1-2 transition can again be fit using Eq.~\ref{eqn:gibbs}, with probabilities as before for the 1-electron states and $p_{S/T}(B_\parallel, T) = (1+ e^{\mp \frac{g\mu_B B_\parallel - \Delta_{ST}}{k_B T}})^{-1}$ for the 2-electron singlet-triplet states, where $\Delta_{ST}$ is the singlet-triplet splitting at zero field. The fit allows for an offset from $dS=0$, away from the degenerate points, to compensate for non-linearities in the charge sensor. From the fit we find the entropies at the two-fold degenerate points, $B=0$ and \SI{9}{\tesla}, are within 5\% of the expected values, $\mp k_B \ln{2}$. The extracted value of $g = 0.39$ from the low-field peak agrees well with the fit in Fig.~\ref{fig:fig3}a. At the high-field singlet-triplet degeneracy we find $g = 0.59$. This unexpected g-factor is explained by a shifting of the $T_{0}$ state with magnetic field, as seen in Fig.~\ref{fig:fig4}f and previous work [CITE]. The high quality of this Gibbs entropy fit confirms $dS$ can be related directly to a tunable degeneracy in the system.

@@

The conductance of the charge sensor was measured at a DC voltage bias of \SI{200}{\micro\volt}, with the current read out directly or fed into a lockin amplifier.  The temperature of the reservoir could be raised above $T_{MC}$ (to $T_{MC} + dT$) using a current biased through the left (heater) QPC. A calibration for $dT$ in this mode, using DC current bias $\alpha$ as measured in Fig.~\ref{fig:fig2}b, can be seen in Fig.~\ref{fig:fig2}c. Applying AC current at $f_{Heat} =$ \SI{48.7}{\hertz} yields oscillations in $T_{res}$ at $2f_{heat}$; to leading order $dT_{res} \sim P \sim \cos(2 \times (2 \pi f_{heat}) \times t)$. Variations in $g_{sens}$ with $dT$ are captured with a lockin amplifier measuring $g_{sens}$ oscillations at at the second harmonic of $f_{heat}$.

%\acknowledgments
% Everyone I might acknowledge is currently an author. Consider moving someone here.

>>>>

% add citations for AFM papers (x2) for degeneracy measurements (vs T) and Beenakker (x1) for Gamma eqn

%%%% outstanding mysteries %%%%%
% error bar on 0-1 peak height (mention in text)
% can I cite someone as to why \Delta_ST changes with field

@@

\begin{abstract}

Measuring the entropy of an electronic state is a powerful tool for identifying its underlying microscopic character.  Such measurements are typically based on bulk properties, such as heat capacity, that are straightforward to observe in macroscopic samples but exceedingly difficult to access in mesoscopic systems that may consist of just a few electrons. Taking advantage of a well-known Maxwell relation, we realize a protocol for entropy-to-charge conversation in a gate-defined GaAs quantum dot that enables an entropy measurement of the first three quantum states in to the dot. The entropy of a single spin ($k_B \ln{2}$) is measured within 8\% accuracy, as is the entropy arising at the magnetic field-driven singlet-triplet crossing for two electrons.

\end{abstract}

@@

%%%%%%% intro material %%%%%%%%%
The thermodynamic properties of electronic systems can offer important insights into the nature of their ground states, and probe exotic quasiparticles that may emerge due to interactions or non-trivial topology.  Systems that are difficult to clearly identify through standard conductance measurements may be studied more in depth if a thermodynamic measurement can be made. For example, the purported non-Abelian exchange statistics of Moore-Read quasiparticles in the $\nu = \frac{5}{2}$ fractional quantum Hall state, or of Majorana quasiparticles in a topological superconductor, are exceedingly difficult to identify from conductance signatures. However, an entropy measurement could clearly distinguish Abelian from non-Abelian quasiparticles \cite{Cooper2009, Smirnov2015}.  Similarly, the two-channel Kondo state that is believed to arise in carefully tuned GaAs devices has so far been identified through a particular temperature dependence in the device conductance \cite{Potok2007}. A more direct test for the two-channel Kondo state would be a confirmation of its entropy ($\frac{1}{2} k_B \ln{2}$), that should remain down to arbitrarily low temperatures \cite{Alkurtass2016}.

%%% figure 1 %%%
\begin{figure}[!]
        \includegraphics[width=1.0\columnwidth]{../figures/figure_1_no-annotation.pdf}
        \caption{\label{fig:fig1}(a) Scanning electron micrograph of a device similar to the one measured, showing electrostatic gates (blue) used to define a circuit in the 2D electron gas. Grey gates are unused (grounded). Squares show the locations of ohmic contacts to the 2DEG. An electron reservoir is formed between the parallel gates running vertically down the image. The reservoir can be heated by driving current through the quantum point contact on the left. A 200nm diameter quantum dot is coupled to the right side of the channel. Electrons tunneling between the reservoir and dot are measured with a capacitively-coupled charge sensor. (b) and (c) Charge sensor signal for a single transition from $N-1 \rightarrow N$ electrons at two temperatures ($T_{Red} > T_{Blue}$) showing the two possible cases for $\frac{\partial \mu}{\partial T}$.}
\end{figure}

@@

In order to accomplish this entropy to charge conversion, we make use of the Maxwell relation
%
\begin{align}
\label{eqn:max}
        \left(\frac{\partial \mu}{\partial T}\right)_{p,N} &= -\left(\frac{\partial S}{\partial N}\right)_{p,T}
\end{align}
%
that connects changes in entropy, particle number, and temperature ($S$, $N$, and $T$, respectively) to changes in the chemical potential, $\mu$, a quantity that is simple to measure and control. The fixed pressure condition of Eq.~\ref{eqn:max} is effectively met by working well below the Fermi temperature of the 2DEG, $T_F \sim$\SI{5}{\kelvin}, where degeneracy pressure dominates \cite{Landau1969}.

The Maxwell relation in Eq.~\ref{eqn:max} forms the basis of two exciting theoretical proposals to measure the entropy derivative, $\frac{\partial S}{\partial N}$, of the $\nu = \frac{5}{2}$ state by detecting changes in the quasiparticle chemical potential with temperature \cite{Cooper2009,Ben-Shach2013}. Unlike the $\nu = \frac{5}{2}$ quasiparticles described in the theoretical work, the few-electron quantum dot states investigated here have an entropy that is well-understood from the spin degree of freedom \cite{Tarucha1996, Ciorga2000, Duncan2000, Lindemann2002, Potok2003, Hofmann2016}. Applying the language of quantum dots to Eq.~\ref{eqn:max}, the shift in electrochemical potential, $\delta \mu$, with temperature is proportional to the entropy difference between the $N-1$ to $N$ quantum dot ground states ($\Delta S_{N-1\rightarrow N}$ for $\Delta N=1$). This shift, $\delta \mu$ as a function of $\delta T$, is the quantity probed directly in the experiment described below. From these measurements the entropies of the first three quantum dot levels are built up, demonstrating a direct entropy measurement is possible on a few-particle system. In the future, the method can be utilized to probe the nature of more exotic systems, such as topologically non-trivial Majorana and other non-Abelian states.

The mesoscopic circuit shown in Fig.~\ref{fig:fig1}a uses electrostatic gates to realize an electron reservoir (central region; hereafter ``reservoir") in thermal and diffusive equilibrium with a few-electron quantum dot (``dot") coupled to its right side.  The occupation of the dot is measured using an adjacent quantum point contact (``charge sensor")\cite{Staring2007, Thierschmann2015}.  The $N$th is added to an $N-1$ electron dot when the electrochemical potential, $\mu_{N}$, drops below the Fermi level of the reservoir, $E_F$. Figure 1b illustrates a such a transition \textemdash a step in the charge sensor conductance as a function of plunger gate voltage, $V_G$, which tunes $\mu_{N}$ \textemdash thermally broadened by the reservoir temperature, $T_{res}$.  The gate voltage corresponding to midpoint of the transition, $V_{mid}$, marks the electrochemical potential at which the probabilities of finding $N-1$ or $N$ electrons on the dot are equal and the $N$th level is occupied half the time.

The relation between entropy and a temperature-induced shift in $\mu_{N})$ (with respect to $E_F$) can be understood in the language of detailed balance.  The $N$th electron tunnels back and forth between an $N-1$-electron dot and reservoir as long as there are available states in both the dot and reservoir; true when $\mu_{N})$ is within thermal broadening of $E_F$.  The tunnel rates into and out of the dot, $\Gamma_{in}=\Gamma_{N-1\rightarrow N}$ and $\Gamma_{out}=\Gamma_{N\rightarrow N-1}$, depend on degeneracy of the $N-1$ and $N$ ground states.  Consider the filling of the first electron state, which has a single, unpaired spin-$\frac{1}{2}$ and, therefore, a degeneracy of 2. An electron tunneling into the dot has both spin-up and -down states available.  Tunneling out, the electron (whether spin-up or -down) must tunnel into reservoir states with the same spin, that is, only half of the reservoir states are available. As a result, $\Gamma_{in} = 2\Gamma_{out}$ when $\mu_{N}=E_F$. Extending this logic to a general case gives the result derived from detailed balance \cite{Gustavsson2009}, 
%
\begin{align}
	\Gamma_{in} &=  g_{N} \Gamma f(E_F - \mu_{N}) \nonumber \\
	\Gamma_{out} &= g_{N-1} \Gamma [1 - f(E_F - \mu_{N})] \label{eqn:rates}
\end{align}
%
where $\Gamma$ is a coupling constant, $f$ is the Fermi function, and $g_{N(N-1)}$ is the degeneracy of the $N(N-1)$ electron state.

The midpoint of the transition between the $N-1$ and $N$ electron states ($V_{mid}$), occurs when $\Gamma_{in} = \Gamma_{out}$. From Eq.~\ref{eqn:rates}, $\Gamma_{in} = \Gamma_{out}$ corresponds to $\mu_{N} = E_F$ when $g_{N-1}=g_{N}$. When $g_{N-1}\neq g_{N}$, $\mu_{N}$ shifts away from $E_F$ by an amount that depends on the broadening of the Fermi function, and therefore on $T$.   A measurement of how $V_{mid}$ shifts with T allows the ratio $g_{N}/g_{N-1}$ to be determined, which itself is related to $\Delta S_{N-1\rightarrow N}$ through the Boltzmann entropy, $S_{N}=k_{B} \ln{g_N}$.  The relationship between degeneracies and tunnel rates was recently explored in a time-resolved measurement \cite{Hofmann2016}. The thermal technique presented in this work extends the measurement of entropy (degeneracy) to a wider set of applications, where the tunneling process may not be convenient, or even possible, to observe in real-time.

@@

%%% figure 2 %%%
\begin{figure}[!]
        \includegraphics[width=1.0\columnwidth]{../figures/figure_2.pdf}
        \caption{\label{fig:fig2}(a) Charge sensor data for $N=0 \rightarrow 1$ at two temperatures set by DC current through the QPC heater. (b) Lock-in measurement of $dg_{sens}$ with $\delta T \sim 60mK$. Fits to $dg_{sens}$ are shown with $\delta$ as a free parameter (solid) and $\delta=0$ (dashed) (c) Transition width, $\Theta$, as a function of mixing chamber temperature. The highlighted region shows roughly where tunnel rates are determined only by temperature broadening. The lever arm $\alpha$ is calculated by fitting a straight line to this region. (d) $\Theta$ as a function of DC current through the QPC heater. Fit is used to convert between $I_{heat}$ and $\delta T$, with $R_{QPC} = $ \SI{20}{\kilo\ohm} (e) Change in entropy as a function of AC current through the QPC heater. Top axis shows corresponding $\delta T$. Right axis shows shift in $\delta V_{mid}$ per unit temperature.}
\end{figure}

@@

%%% figs 2a and b %%%
The addition of the first electron to the dot was marked by a decrease in $g_{sens}$ that is well fit by the expected form for a thermally-broadened transition (Fig.~\ref{fig:fig2}a):
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_{mid}(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_{mid}(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $G_0$ quantifies the sensor sensitivity, $\Theta = \frac{k_B T}{\alpha e}$ is the thermal broadening expressed in units of gate voltage, $\alpha$ is the lever arm that converts $V_G$ to $\mu_{N}$, and $G_1$ is determined by the cross capacitance between the charge sensor and plunger gate. The dependence of $V_{mid}$ on $\Theta$ captures any shift in $\mu_{N}$ due to a change in entropy across the transition.

The transition width, $\Theta$, extracted from fits to Eq.~\ref{eqn:g} follows the mixing chamber temperature, $T_{MC}$, to \SI{<100}{\milli\kelvin} (Fig.~\ref{fig:fig2}b). The assumption of thermally broadened transitions greatly simplifies the data analysis, so $T_{MC}$ is fixed at \SI{100}{\milli\kelvin} for most of this work with Joule heating giving reservoir temperatures up to \SI{250}{\milli\kelvin}. Measurements of $\Delta S$ are made at $\delta T_{res} = $ \SI{60}{\milli\kelvin} such that no orbital levels above the ground state were excited.

%%% figs 2c and d%%%
The data in Fig.~\ref{fig:fig2}c, and corresponding fits, offer an example of how $\Delta S_{0\rightarrow 1}$ was measured across the $0 \rightarrow 1$ transition. A comparison of the 100 and 160 mK data in Fig.~\ref{fig:fig2}a shows that $g_{sens}$ drops as the temperature rises before the midpoint of the transition ($V_G-V_{mid}<0$), whereas the conductance is higher in the warmer curve for $V_G-V_{mid}>0$.   Oscillating the reservoir temperature by $\delta T_{res}$ using an AC current, $I^{AC}_{Heat}$, while measuring oscillations in $g_{sens}$ with a lock in amplifier (see Methods) yields the characteristic peak-dip structure seen in Fig.~\ref{fig:fig2}c.  The expected lineshape of such a curve is given by the derivative of Eq.~\ref{eqn:g} with respect to temperature, expressed as $dg_{sens}$ due to a temperature change $\delta T$:
%
\begin{align}
\label{eqn:dg}
        dg_{sens}(V_G, \Theta) &\propto \delta T \left[ \frac{V_G-V_{0}(\Theta)}{2\Theta} +\delta \right]\times \\
        				      &\quad\cosh^{-2}\left(\frac{V_G-V_{0}(\Theta)}{2\Theta}\right) + const. \nonumber
\end{align}
%
where $\delta=\frac{\partial V_{mid}}{\partial \Theta}$. Using the lever arm to convert $V_{mid}$ to electrochemical potential, $\delta$ is expressed as the change in entropy (scaled by $k_B$) through Eq.~\ref{eqn:max}:
%
\begin{align}
\label{eqn:delta}
        \delta &= \frac{\partial V_{mid}}{\partial \Theta} = 
        \frac{1}{k_B} \frac{\partial \mu}{\partial T} = 
        -\frac{1}{k_B} \Delta S_{N-1\rightarrow N}
\end{align}
%

From Eqs.~\ref{eqn:dg} and \ref{eqn:delta}, $dg_{sens}(V_G)$ would be perfectly antisymmetric around $V_{mid}$ if $\Delta S_{N-1\rightarrow N}$ were zero.  Fits to the data in Fig.~\ref{fig:fig2}c illustrate clearly that $\Delta S_{0\rightarrow 1}=0$ is inconsistent with the data;  the best fit is found for $\Delta S_{0\rightarrow 1}=0.71\times k_B$.  Indeed, this is what would be expected for the $0 \rightarrow 1$ electron transition at zero magnetic field: $S_0$ for the empty state should be $k_B \ln{1}=0$ whereas for the spin degenerate 1-electron state one expects $S_1=k_B\ln{2}$, giving $\Delta S_{0\rightarrow 1}\equiv S_1 = S_0 =k_B\ln{2}$. Fig.~\ref{fig:fig2}e shows that $\Delta S$,  determined from fits to data analogous to Fig.~\ref{fig:fig2}c, remains constant over a range of $\delta T$ ($I^{AC}_{Heat}$).  Although the extracted $\delta$ can be converted to gate-voltage-per-temperature units using a value for $\alpha$ measured separately (right axis of Fig.~\ref{fig:fig2}e), the entropy itself, $\Delta S$, is determined directly from the fit to $dg_{sens}$ without any additional calibration.

@@

Confirmation that the measured $\Delta S$ derives from spin degeneracy can be seen in its evolution with in-plane magnetic field, $B_\parallel$. Figure \ref{fig:fig3}a compares $\Delta S(B_\parallel)$ for the 0-1 and 2-3 electron transitions.  The 2-electron ground state is a spin singlet, while the 3-electron ground state adds an additional unpaired electron, so both $\Delta S_{0 \rightarrow 1}$ and $\Delta S_{2 \rightarrow 3}$ correspond to transitions from total spin zero $(S_0=S_2=0)$ to total spin one-half.One expects identical field-dependent behaviors for $\Delta S_{0\rightarrow 1}$ and $\Delta S_{2\rightarrow 3}$ as long as $g \mu_{B} B_\parallel$ is less than the excited state energies.

@@

The entropy measurement yields $\Delta S(B=0) = (0.933 \pm 000) k_B \ln{2}$, which agrees quantitatively with the expected $k_{B} \ln{2}$, collapsing to zero at high field where spin degeneracy is broken.  At a qualitative level, Figs.~\ref{fig:fig3}b and c show the high-field crossover to a perfectly antisymmetric lineshape for $dg_{sens}(V_G)$; clearly indicating $\delta$ from Eqn.~\ref{eqn:dg} going to zero, corresponding to no entropy change between the two charge states. Using a $T=T_{res}$ of \SI{130}{\milli\kelvin} that is midway between the low ($T_{MC}$) and high ($T_{MC}+\delta T$) points of the temperature oscillation yields $g=-0.39$ and $g=-0.40$ for the $0\rightarrow 1$ and $2\rightarrow 3$ transitions, respectively. This is consistent with the value $g=-0.44$ expected for GaAs and measured via bias spectroscopy in Fig.~\ref{fig:fig3}e.

@@

The field-dependent entropy measurement for the 1-2 transition can again be fit using Eq.~\ref{eqn:gibbs}, with probabilities as before for the 1-electron states and $p_{S/T}(B_\parallel, T) = (1+ e^{\mp \frac{g\mu_B B_\parallel - \Delta_{ST}}{k_B T}})^{-1}$ for the 2-electron singlet-triplet states, where $\Delta_{ST}$ is the singlet-triplet splitting at zero field. The fit allows for an offset from $\Delta S=0$, away from the degenerate points, to compensate for non-linearities in the charge sensor. From the fit we find the entropies at the two-fold degenerate points, $B=0$ and \SI{9}{\tesla}, are within 5\% of the expected values, $\mp k_B \ln{2}$. The extracted value of $g = 0.39$ from the low-field peak agrees well with the fit in Fig.~\ref{fig:fig3}a. At the high-field singlet-triplet degeneracy we find $g = 0.59$. This unexpected g-factor is explained by a shifting of the $T_{0}$ state with magnetic field, as seen in Fig.~\ref{fig:fig4}f and previous work [CITE]. The high quality of this Gibbs entropy fit confirms $\Delta S$ can be related directly to a tunable degeneracy in the system.

@@

The conductance of the charge sensor was measured at a DC voltage bias of \SI{200}{\micro\volt}, with the current read out directly or fed into a lockin amplifier.  The temperature of the reservoir could be raised above $T_{MC}$ (to $T_{MC} + \delta T$) using a current biased through the left (heater) QPC. A calibration for $\delta T$ in this mode, using DC current bias $\alpha$ as measured in Fig.~\ref{fig:fig2}b, can be seen in Fig.~\ref{fig:fig2}c. Applying AC current at $f_{Heat} =$ \SI{48.7}{\hertz} yields oscillations in $T_{res}$ at $2f_{heat}$; to leading order $\delta T_{res} \sim P \sim \cos(2 \times (2 \pi f_{heat}) \times t)$. Variations in $g_{sens}$ with $dT$ are captured with a lockin amplifier measuring $g_{sens}$ oscillations at at the second harmonic of $f_{heat}$.

%\acknowledgments
